Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-27 Chapter-9/9-Uncertainty.tex
s

\subsection{$E_{loss}$}
We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Monte Carlo using the DDMC pion sample, since this quantity is not  measurable directly on data. We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample. The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determine the trajectory of a particle and whether or not it will hit the halo paddle. Thus, to

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss60A.png}
\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).}
\label{fig:ELoss60A}
\end{figure}

@@

The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.

\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
Back of the envelope material budget calculation:}
\begin{table}[h!]
\centering
\caption{Back of the envelope calculation}
\label{my-label}
\begin{tabular}{|l|l|l|l|}
\hline
dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
Total                                               &                                      &                & 12.4                  \\ \hline
\end{tabular}
\end{table}

\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}

>>>>

\subsection{$E_{loss}$}
We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Monte Carlo using the DDMC pion sample, since this quantity is not  measurable directly on data. We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample. The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle. In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation. These distributions can be separated drawing a line in this position-momentum space. 
We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line. We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation. Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
We apply the same classifier on data. We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".

%We use the separation of these two distributions to decide what the energy loss for each event on data. 
%Thus,  we assign the value for energy loss is used in the data.

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss60A.png}
\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right). The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.  }
\label{fig:ELoss60A}
\end{figure}

@@

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-9/Images/PXVsX60A.png}
\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs. The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo. The form of the classifier is overlaid to both plots (red line).}
\label{fig:PxVsXTrue}
\end{figure}

%\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
%The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
%Back of the envelope material budget calculation:}
%\begin{table}[h!]
%\centering
%\caption{Back of the envelope calculation}
%\label{my-label}
%\begin{tabular}{|l|l|l|l|}
%\hline
%dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
%1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
%1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
%1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
%1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
%1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
%Total                                               &                                      &                & 12.4                  \\ \hline
%\end{tabular}
%\end{table}
